Consistent Polyhedral Surrogates via Embedding

### 2019-10-24 notes/top-k.tex
top-k comments

%We use $\odot$ to denote the Hadamard (element-wise) vector product.
Let $\min(\vec w, \vec v)$ denote the element-wise minimum of vectors $\vec w, \vec v \in \reals^d$.

@@

\begin{lemma}\label{lem:extra-elts-zero}
	For all $y \in \Y$, there is a minimizer $u$ of $L^{top-k}(\cdot)_y$ satisfying Assumption~\ref{assum:ordered-u} such that $u_{k+1} = \ldots = u_{n-1} = u_n = 0$.
\end{lemma}
\begin{proof}
	By Lemma~\ref{lem:invar-ones}, we can fix $u_{k+1} = 0$.
	With Assumption~\ref{assum:ordered-u}, we then have $u_i \leq 0$ for all $i \geq k+1$.
	Consider $u'$ such that $u'_i = u_i$ for all $i$ so that $u_i \geq 0$, and $u'_i = 0$ otherwise.

@@

\section{Re-writing the loss}
The following Lemma allows us to consider the loss without the $(\cdot)_+$ operator.

@@

With Assumptions~\ref{assum:ordered-u}, \ref{assum:nonneg}, \ref{assum:top-threshold} and Lemma~\ref{lem:get-rid-max-op}, we can know re-write the loss for each outcome in a simplified form.
\begin{lemma}\label{lem:new-form-surrogate}
	For $u$ satisfying Assumptions~\ref{assum:ordered-u},~\ref{assum:nonneg}, and~\ref{assum:top-threshold}, the top-$k$ surrogate can be re-written:
\begin{align*} 
L^{top-k}(u)_y  &= 1 - u_y + \bar u - \frac{1}{k}(\min(u_y, 1))~.~
\end{align*}
\end{lemma}
\begin{proof}
	First, consider
\begin{align*}
	L^{top-k}(u)_y &= \left(1 -u_y + \frac{1}{k} \sum_{i=1}^k ((u - e_y)_{[i]} \right)_+\\
	&= 1 -u_y + \frac{1}{k} \sum_{i=1}^k ((u - e_y)_{[i]}~,
	\end{align*}
	by Lemma~\ref{lem:get-rid-max-op}.

@@

\begin{lemma}\label{lem:linear-pieces}
	For each element $u_i$ of $u$, the expected loss $\inprod{p}{L^{top-k}(u)}$ is linear on $[0,1]$ and $[1, \bar u_{-i} + 1]$ for all $p \in \simplex$.
	\jessie{Not sure if this is the right statement.}	
\end{lemma}
\begin{proof}
	For all $p \in \simplex$ and $u$ satisfying Assumptions~\ref{assum:ordered-u},~\ref{assum:nonneg}, and~\ref{assum:top-threshold}, we know that 
	\begin{equation}
	\inprod{p}{L^{top-k}(u)} = 1 - \bar u + \inprod{p}{u + \frac 1 k \min{u,\ones}}~.~
	\end{equation}
	For any $p$, if $u_i \in [0,1]$, we then have 
	\begin{align*}
	\inprod{p}{L^{top-k}(u)} &= 1 + \bar u - \inprod{p}{u + \frac 1 k \min{u,\ones}}\\
	 &= 1 + \bar u - \sum_{y \neq i}p_y (u_y + \frac 1 k \min(u_y, 1)) - p_i \frac {k+1} k u_i~,~
	\end{align*}
	which is linear on $u_i$.

@@

\begin{theorem}\label{thm:finite-set}
	Consider the loss $L^{top-k} : \reals^n \times \Y \to \reals_+$ given in Equation~\ref{eq:top-k-surrogate}.  
	For all $p \in \simplex$, there is a report in the following report set $\R$ minimizing $\inprod{p}{L^{top-k}(\cdot)}$.
	\begin{align*}
	\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}
	\end{align*}
	\jessie{Check the $|H|+|M| \leq k$ condition. Worst case, the set $\R$ is still finite without it.}
\end{theorem}
\begin{proof}[Proof Sketch.]
%	The above work tells us that when $p_y > \frac 1 {k+1}$, it is optimal to set $u_y > 0$, and otherwise, it is optimal to set $u_y = 0$.
	Since $L^{top-k}$ is linear on $[0,1]$ and $[1, c]$ for some $c \geq 1$ and all dimensions $i \in [n]$, we know that the optimum is on the boundary of such reports, so when we bump $u_y$ up to some number greater than $0$, it should either be one $1$ or $c$ to be an optimal report.
	$H$ represents the elements of the report that are bumped up to $c$, and $M$ represents the elements of the report that are bumped up to $1$.
	We know that the size of $H \cup M$ must be less than or equal to $k$ by \jessie{???}, and $H \cap M$ is empty by each element of $u$ being real-valued, so for $c > 1$, $u_i$ cannot simultaneously be equal to $c$ and $1$, hence the restrictions on $\R$.
	Note that if $c = 1$, then we can say $H = \emptyset$.

>>>>

%We use $\odot$ to denote the Hadamard (element-wise) vector product.
Let $\min(\vec w, \vec v) \in \reals^d$ denote the element-wise minimum of vectors $\vec w, \vec v \in \reals^d$.

@@

\begin{lemma}\label{lem:extra-elts-zero}
  \raf{This statement seems off.  I think you want to say that for all $u\in\reals^d$, there is a $u'$ satisfying assumption~\ref{assum:nonneg} such that $L(u')_y \leq L(u)_y$ for all $y$ (pardon lazy notation).}
	For all $y \in \Y$, there is a minimizer $u$ of $L^{top-k}(\cdot)_y$ satisfying Assumption~\ref{assum:ordered-u} such that $u_{k+1} = \ldots = u_{n-1} = u_n = 0$.
\end{lemma}
\begin{proof}
	By Lemma~\ref{lem:invar-ones}, we can fix $u_{k+1} = 0$.
	With Assumption~\ref{assum:ordered-u}, we then have $u_i \leq 0$ for all $i \geq k+1$.
	Consider $u'$ such that $u'_i = u_i$ for all $i$ so that $u_i \geq 0$, and $u'_i = 0$ otherwise.

@@

\section{Re-writing the loss}
The following Lemma allows us to consider the loss without the $(\cdot)_+$ operator.
\raf{I think it might be handy to introduce the set $\U$ here to be the points satisfying all three assumptions --- maybe explicitly write the constraints and nod to the assumptions.  Then we can say $u\in\U$ and forget the assumptions going forward.}

@@

With Assumptions~\ref{assum:ordered-u}, \ref{assum:nonneg}, \ref{assum:top-threshold} and Lemma~\ref{lem:get-rid-max-op}, we can know re-write the loss for each outcome in a simplified form.
\raf{Note: some inconsistency in the way you use this $\min$ operator and denote $\ones$; see below too}
\begin{lemma}\label{lem:new-form-surrogate}
	For $u$ satisfying Assumptions~\ref{assum:ordered-u},~\ref{assum:nonneg}, and~\ref{assum:top-threshold}, the top-$k$ surrogate can be re-written:
\begin{align*} 
L^{top-k}(u)_y  &= 1 - u_y + \bar u - \frac{1}{k}(\min(u_y, 1))~.~
\end{align*}
\end{lemma}
\begin{proof}
	First, consider
\begin{align*}
	L^{top-k}(u)_y &= \left(1 -u_y + \frac{1}{k} \sum_{i=1}^k ((u - e_y)_{[i]} \right)_+\\
	&= 1 -u_y + \frac{1}{k} \sum_{i=1}^k ((u - e_y)_{[i]}~,
	\end{align*}
	by Lemma~\ref{lem:get-rid-max-op}.

@@

\begin{lemma}\label{lem:linear-pieces}
	For each element $u_i$ of $u$, the expected loss $\inprod{p}{L^{top-k}(u)}$ is linear on $[0,1]$ and $[1, \bar u_{-i} + 1]$ for all $p \in \simplex$.
	\jessie{Not sure if this is the right statement.}
    \raf{Not sure either.  One thing to be careful about: you either have to make sure that the intervals in question yield points in the set $\U$, or refer to the original form of the loss.  Because the form of the loss you are using only is shown to hold when $u\in\U$.  Hopefully that made sense.}
    \raf{One thing that would help: let's visualize $\U$ for some value of $k,n$ to get some intuition for what this set is.  Maybe $k=2$ and $n=3$, but only plot the first 2 coordinates?}
\end{lemma}
\begin{proof}
	For all $p \in \simplex$ and $u$ satisfying Assumptions~\ref{assum:ordered-u},~\ref{assum:nonneg}, and~\ref{assum:top-threshold}, we know that 
	\begin{equation}
	\inprod{p}{L^{top-k}(u)} = 1 - \bar u + \inprod{p}{u + \frac 1 k \min{u,\ones}}~.~
	\end{equation}
	For any $p$, if $u_i \in [0,1]$, we then have 
	\begin{align*}
	\inprod{p}{L^{top-k}(u)} &= 1 + \bar u - \inprod{p}{u + \frac 1 k \min{u,\ones}}\\
	 &= 1 + \bar u - \sum_{y \neq i}p_y (u_y + \frac 1 k \min(u_y, 1)) - p_i \frac {k+1} k u_i~,~
	\end{align*}
	which is linear on $u_i$.

@@

\begin{theorem}\label{thm:finite-set}
	Consider the loss $L^{top-k} : \reals^n \times \Y \to \reals_+$ given in Equation~\ref{eq:top-k-surrogate}.  
	For all $p \in \simplex$, there is a report in the following report set $\R$ minimizing $\inprod{p}{L^{top-k}(\cdot)}$.
	\begin{align*}
	\R := \left\{ \frac{|M| + k -1}{k - |H|} \mathbf{1}_H + \mathbf{1}_M : H, M \subset [n], H\cap M = \emptyset, |H| + |M| \leq k \right\}
	\end{align*}
	\jessie{Check the $|H|+|M| \leq k$ condition. Worst case, the set $\R$ is still finite without it.}\raf{follows from assump~\ref{assum:nonneg}, no?}
\end{theorem}
\begin{proof}[Proof Sketch.]
%	The above work tells us that when $p_y > \frac 1 {k+1}$, it is optimal to set $u_y > 0$, and otherwise, it is optimal to set $u_y = 0$.
	Since $L^{top-k}$ is linear on $[0,1]$ and $[1, c]$ for some $c \geq 1$ and all dimensions $i \in [n]$, we know that the optimum is on the boundary of such reports, so when we bump $u_y$ up to some number greater than $0$, it should either be one $1$ or $c$ to be an optimal report.
	$H$ represents the elements of the report that are bumped up to $c$, and $M$ represents the elements of the report that are bumped up to $1$.
	We know that the size of $H \cup M$ must be less than or equal to $k$ by \jessie{???}, and $H \cap M$ is empty by each element of $u$ being real-valued, so for $c > 1$, $u_i$ cannot simultaneously be equal to $c$ and $1$, hence the restrictions on $\R$.
	Note that if $c = 1$, then we can say $H = \emptyset$.
